---
title: "BIOS 699 Project 1 Analysis"
author: "Mengbing Li"
date: "January 15, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mengbing/Box Sync/2019 Winter/BIOS699_BiostatisticsInvestigations/project1')

library(dplyr)
library(tidyr)
library(data.table)
library(readr)
library(knitr)
library(ggplot2)
library(lme4)
# library(lmerTest)
```

```{r}
capacity <- fread("thedata.csv")
colnames(capacity) <- tolower(colnames(capacity))
colnames(capacity)[3] <- "age"

driver <- fread("ElderlyDriverData.csv")
colnames(driver) <- tolower(colnames(driver))
colnames(driver)[33:34] <- c("ingresstime", "egresstime")
colnames(driver)[which(colnames(driver)=="subjective.rating..scale.from.1.to.50.")] <- "subjectiveRating"
colnames(driver)[which(colnames(driver)=="age..yrs.")] <- "age"
```



```{r}
mergedData <- merge(x = driver, y = capacity,
                    by = c("subject", "gender", "age"))
colnames(mergedData) <- gsub("\\.", "", colnames(mergedData))

# extract door condition
mergedData$doorCondition <- with(mergedData, ifelse(grepl("Open", vehicleanddoorcondition), "open", "half-open"))
# extract vehicle type
mergedData$vehicle <- parse_number(mergedData$vehicleanddoorcondition)
mergedData$vehicle <- as.factor(mergedData$vehicle)
  
# obtain averaged mobility measurements --------------------
# time up and go
mergedData$tug_mean <- apply(mergedData[, c("tug1", "tug2")], 1, function(x) mean(na.omit(x)))

# one leg balance time
olb_names <- grep("olb", colnames(mergedData), value = TRUE)
mergedData$olb_mean <- apply(mergedData[, olb_names, with = FALSE], 1, function(x) mean(na.omit(x)))

plot(mergedData$tug_mean, mergedData$olb_mean)
cor(mergedData$tug_mean, mergedData$olb_mean, use="complete.obs")
#-0.5512389. moderately strong correlation

# Strength measurements ------------------------------------
kneeext_names <- grep("kneeext", colnames(mergedData), value = TRUE)
mergedData$kneeext_mean <- apply(mergedData[, kneeext_names, with = FALSE], 1, function(x) mean(na.omit(x)))

hipabd_names <- grep("hipabd", colnames(mergedData), value = TRUE)
mergedData$hipabd_mean <- apply(mergedData[, hipabd_names, with = FALSE], 1, function(x) mean(na.omit(x)))

hiprtd_names <- grep("hiprtd", colnames(mergedData), value = TRUE)
mergedData$hiprtd_mean <- apply(mergedData[, hiprtd_names, with = FALSE], 1, function(x) mean(na.omit(x)))

grip_names <- grep("grip", colnames(mergedData), value = TRUE)
mergedData$grip_mean <- apply(mergedData[, grip_names, with = FALSE], 1, function(x) mean(na.omit(x)))


# Cognitive measurements -----------------------------------
hiprtd_names <- grep("hiprtd", colnames(mergedData), value = TRUE)
mergedData$hiprtd_mean <- apply(mergedData[, hiprtd_names, with = FALSE], 1, function(x) mean(na.omit(x)))

# Scatterplot Matrices for reaction times
my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")  
pairs(~ srt + rrt + accuracy + offaccuracy, 
      data = mergedData, 
      pch = 19,  cex = 0.5,
      col = my_cols[as.numeric(factor(mergedData$group))],
      main="Simple Scatterplot Matrix")
## accuracy and offaccuracy are linearly associated
cor(capacity$srt, capacity$rrt)

# calculate BMI
mergedData$BMI <- with(mergedData, weightkg/(staturewithoutshoesmm/1000)^2)
```



## 1. Does egress strategy and mobility impairments impact egress time?

```{r}
model0 <- lmer(data = mergedData, egresstime ~ egressstrategy + tug_mean + olb_mean + (1|subject))
summary(model0)
```

```{r}
anova(model0)
```

```{r}
qqnorm(resid(model0))
qqline(resid(model0))
shapiro.test(resid(model0))
```


```{r}
model0.2 <- lmer(data = mergedData, egresstime ~ egressstrategy + tug_mean + (1|subject))
summary(model0.2)
```

Hence, egress stragety is not significantly associated with egress time. Tug mean is significant.



## 2. Identify human capacities that affect egress time

```{r}
model1 <- lmer(data = mergedData, log(egresstime) ~ egressstrategy + tug_mean + olb_mean + age + gender + BMI + doorCondition + vehicle + group + kneeext_mean + hipabd_mean + hiprtd_mean + grip_mean + mmse + trailsb + srt + rrt + accuracy + (1|subject))
summary(model1)
```



It turns out that egress stragety is strong affected by doorCondition. We will keep egress stragety for scientific interest.



The largest p values come from trailsb, grip_mean, srt, hipabd_mean, hiprtd_mean, egressstrategy. We drop these variables and see whether the reduced model is sufficient.

```{r}
model2 <- lmer(data = mergedData, egresstime ~ egressstrategy + tug_mean + olb_mean + age + gender + BMI + doorCondition + vehicle + group + kneeext_mean + mmse + rrt + accuracy + (1|subject))
(s2 <- summary(model2))
(1-pt(abs(s2$coefficients[,3]), 624))*2
```


All coefficients except for egress stragety do not change much after dropping the variables. So we are fine with dropping the variables.


```{r}
model2.dataNomiss <- na.omit(mergedData[, c("subject", "egressstrategy", "egresstime", "tug_mean", "vehicle", "olb_mean", "age", "gender", "group", "mmse", "BMI", "kneeext_mean", "hipabd_mean", "hiprtd_mean", "grip_mean", "trailsb", "rrt", "srt", "accuracy", "doorCondition")])

model1.nomiss <- lmer(data = model2.dataNomiss, egresstime ~ egressstrategy + tug_mean + olb_mean + age + gender + BMI + doorCondition + vehicle + group + kneeext_mean + hipabd_mean + hiprtd_mean + grip_mean + mmse + trailsb + srt + rrt + accuracy + (1|subject))

model2.nomiss <- lmer(data = model2.dataNomiss, egresstime ~ egressstrategy + tug_mean + olb_mean + age + gender + BMI + doorCondition + group + kneeext_mean + mmse + rrt + accuracy + (1|subject))
# summary(model2.nomiss)
anova(model1.nomiss, model2.nomiss)
```

We do not consider model 1 and model 2 are different. The coefficients do not change much, except for accuracy. So we are fine with dropping the variables. 



Now we continue dropping egress strategy, olb_mean (because of its strong correlation with tug_mean), and mmse.
```{r}
model3 <- lmer(data = mergedData, egresstime ~ tug_mean + age + gender + BMI + group + kneeext_mean + rrt + accuracy + vehicle + doorCondition + (1|subject))
(s3 <- summary(model3))
(1-pt(abs(s3$coefficients[,3]), 637))*2
```


Since some missing values in model 2 are excluded, we need to exclude the same observations in model 3 in order to to model comparison.

```{r}
mergedData.model3 <- na.omit(mergedData[, c("subject", "egressstrategy", "egresstime", "tug_mean", "olb_mean", "age", "gender", "BMI", "group", "kneeext_mean", "mmse", "rrt", "accuracy", "vehicle", "doorCondition")])
model2.nomiss <- lmer(data = mergedData.model3, egresstime ~ egressstrategy + tug_mean + olb_mean + age + gender + BMI + doorCondition + vehicle + group + kneeext_mean + mmse + rrt + accuracy + (1|subject))
model3.nomiss <- lmer(data = mergedData.model3, egresstime ~ tug_mean + age + gender + BMI + group + kneeext_mean + rrt + egressstrategy + vehicle + doorCondition + (1|subject))
anova(model2.nomiss, model3.nomiss)
```




#### Model 4
```{r}
model4 <- lmer(data = mergedData, egresstime ~ tug_mean + gender + BMI + group + kneeext_mean + rrt + vehicle + doorCondition + (1|subject))
(s4 <- summary(model4))
(1-pt(abs(s4$coefficients[,3]), 637))*2
```

```{r}
anova(model3, model4)
```



#### Model 5
```{r}
model5 <- lmer(data = mergedData, log(egresstime) ~ tug_mean + BMI + group + kneeext_mean + vehicle + doorCondition + (1|subject))
(s5 <- summary(model5))
(1-pt(abs(s5$coefficients[,3]), 637))*2
```

```{r}
anova(model4, model5)
```

```{r}
qqnorm(resid(model5))
qqline(resid(model5))
shapiro.test(resid(model5))
```


#### Model 6
```{r}
model6 <- lmer(data = mergedData, egresstime ~ tug_mean + BMI + vehicle + doorCondition + (1|subject))
(s6 <- summary(model6))
(1-pt(abs(s6$coefficients[,3]), 637))*2
```

```{r}
anova(model5, model6)
```


```{r}
qqnorm(resid(model6))
qqline(resid(model6))
shapiro.test(resid(model6))

plot(model6, 
     main = "")
```

Model 6 is the final model.
Calculate model 6 coefficients.
```{r}
s6$coefficients[,1] + 1.96*s6$coefficients[,2]
s6$coefficients[,1] - 1.96*s6$coefficients[,2]
```

```{r}
confint(model6)
```


```{r}
library(MuMIn)
#Determine R2:
r.squaredGLMM(model6) 
```

