---
title: "BIOS 699 Project 5"
author: "Mengbing Li"
date: "April 18, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mengbing/Box Sync/2019 Winter/BIOS699_BiostatisticsInvestigations/project5')

library(dplyr)
library(tidyr)
library(data.table)
library(reshape2)
library(knitr)
library(xlsx)
library(ggplot2)
```

```{r}
capacity <- fread("thedata.csv")
colnames(capacity) <- tolower(colnames(capacity))
colnames(capacity)[3] <- "age"

# driver <- read.xlsx("ElderlyDriverData.xlsx", sheetName = "FeedbackReport_IE 1st Row 2014 ")
# driver[,37:38] <- NULL
# write.table(driver, file = "ElderlyDriverData.csv", quote = FALSE, sep = ",",
#             row.names = FALSE, col.names = TRUE)
driver <- fread("ElderlyDriverData.csv")
colnames(driver) <- tolower(colnames(driver))
colnames(driver)[33:34] <- c("ingresstime", "egresstime")
colnames(driver)[which(colnames(driver)=="subjective.rating..scale.from.1.to.50.")] <- 
  "subjectiveRating"
```



## Exploratory Data Analysis

Things to explore:

1. bar plot of gender distribution

```{r}
counts_gender <- table(capacity$gender)
barplot(counts_gender, main="Gender Distribution", 
   xlab="Gender")
```

2. histogram of age distribution
```{r}
hist(capacity$age, main = "Age distribution", xlab = "Age")

ggplot(data = capacity, aes(x = age, fill = gender)) +
  geom_histogram(bins = 8, alpha = 0.4) +
  labs(title = "Age distribution by gender", x = "Age")

summary(capacity$age)
```


3. gender distribution by group
```{r}
table(capacity$gender, capacity$group)
```

4. age distribution by group
```{r}
ggplot(data = capacity, aes(x = age, fill = group)) +
  geom_histogram(bins = 5, alpha = 0.4) +
  labs(title = "Age distribution by group", x = "Age")
```


5. scatter plot of ingress vs egress time
```{r}
ggplot(data = driver, aes(x = ingresstime, y = egresstime)) +
  geom_point(aes(color = subjectiveRating)) +
  scale_colour_gradient(low = "grey", high = "black") +
  labs(title = "Overall ingress time vs egress time",
       x = "Ingress time (s)", y = "Egress time (s)")
```

There is a positive correlation between ingress time and egress time.

Enlarge the lower-left area to see overlapping points.
```{r}
ggplot(data = driver[driver$ingresstime<10 & driver$egresstime<10,], aes(x = ingresstime, y = egresstime)) +
  geom_point(aes(color = subjectiveRating)) +
  scale_colour_gradient(low = "grey", high = "black") +
  labs(title = "Enlarged plot: Overall ingress time (<10s) vs egress time (<10s)",
       x = "Ingress time (s)", y = "Egress time (s)")
```
People with longer ingress time and egress time seem to have higher subjective rating.



6. ingress and egress time by test
```{r}
ggplot(data = driver, aes(x =vehicle.and.door.condition, y = ingresstime)) +
  geom_boxplot(aes(fill = vehicle.and.door.condition)) +
  geom_point(position = position_dodge(width=0.75), aes(group = vehicle.and.door.condition)) +
  labs(title = "Box plots of ingress time by vehicle and door conditions",
       x = "Vehicle and door conditions", y = "Ingress time")
```

```{r}
driver$vehicle.and.door.condition <- factor(driver$vehicle.and.door.condition, levels = c(paste0("V", 1:7, "_Open"), paste0("V", 1:7, "_Pklt")))

ggplot(data = driver, aes(x = vehicle.and.door.condition, y = egresstime)) +
  geom_boxplot(aes(fill = doorCondition)) +
  # geom_point(position = position_dodge(width=0.75), aes(group = vehicle.and.door.condition, color = factor(vehicle.order))) +
  labs(title = "Box plots of egress time by vehicle and door conditions",
       x = "Vehicle and door conditions", y = "Egress time") +
  theme(legend.position="none",
        plot.title = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 15))
ggsave("boxplot_egresstimeByVehicleDoor.png", height = 6, width = 10)
```


The pattern of ingress time by vehicle condition and egress time by vehicle condition are similar.




Bivariate analyses of ingress time in each vehicle conditions by vehicle order
```{r}
vehicleConditions <- unique(driver$vehicle.and.door.condition)

bivariateIngresstime <- c()
for(name in vehicleConditions){
  fit.byVehicleCondition <- lm(data = driver[vehicle.and.door.condition == name, ], ingresstime ~ vehicle.order)
  fit.byVehicleCondition.summary <- summary(fit.byVehicleCondition)
  bivariateIngresstime <- rbind(bivariateIngresstime, fit.byVehicleCondition.summary$coefficients[2, ])
}

rownames(bivariateIngresstime) <- vehicleConditions
bivariateIngresstime
```


Bivariate analyses of egress time in each vehicle conditions by vehicle order
```{r}
vehicleConditions <- unique(driver$vehicle.and.door.condition)

bivariateEgresstime <- c()
for(name in vehicleConditions){
  fit.byVehicleCondition <- lm(data = driver[vehicle.and.door.condition == name, ], egresstime ~ vehicle.order)
  fit.byVehicleCondition.summary <- summary(fit.byVehicleCondition)
  bivariateEgresstime <- rbind(bivariateEgresstime, fit.byVehicleCondition.summary$coefficients[2, ])
}

rownames(bivariateEgresstime) <- vehicleConditions
bivariateEgresstime
```

From bivariate analysis above, we conclude that the order of testing vehicles does not significantly influence ingress or egress time.



6. cross tabulation of ingress vs egress strageties
```{r}
table(driver$ingress.strategy, driver$egress.strategy)
```

The 317 people whose ingress stragety is normal and egress stragety is two feet are worth investigating.


7. ingress time vs egress time by strageties
```{r}
ggplot(data = driver, aes(x = ingresstime, y = egresstime)) +
  geom_point(aes(color = paste(ingress.strategy, egress.strategy))) +
  # scale_colour_gradient(low = "grey", high = "black") +
  labs(title = "Overall ingress time vs egress time by strategy",
       x = "Ingress time (s)", y = "Egress time (s)")
```

```{r}
ggplot(data = driver[!is.na(ingress.strategy), ], aes(x = ingresstime, fill = ingress.strategy)) +
  geom_histogram(aes(x = ingresstime,y=..density..), alpha = 0.4) + 
  geom_density(aes(x = ingresstime,y=..density..)) +
  labs(title = "Distribution of ingress time by ingress strategy", x = "Ingress Time")
```

```{r}
ggplot(data = driver[!is.na(egress.strategy), ], aes(x = egresstime, fill = egress.strategy)) +
  geom_histogram(aes(x = egresstime,y=..density..), alpha = 0.6) + 
  geom_density(aes(x = egresstime,y=..density..), alpha = 0.4) +
  labs(title = "Distribution of egress time by Egress strategy", x = "Egress Time")
```


9. ingress time and egress time by order
```{r}
ggplot(data = driver[!is.na(egresstime), ], aes(x = egresstime, fill = vehicle.order)) +
  geom_histogram(aes(x = egresstime,y=..density..), alpha = 0.6) + 
  geom_density(aes(x = egresstime,y=..density..), alpha = 0.4) +
  labs(title = "Distribution of egress time by vehicle order", x = "Egress Time")
```




8. spegatti plot of repeated measurements of mobility
```{r}
tugData <- melt(capacity[, c("subject", "tug1", "tug2")], #[, (diag.vars), with = FALSE]
       id.vars = c("subject"),
       measure.vars = c("tug1", "tug2"),
       value.name = "tug",
       variable.name = "measurement",
       na.rm = TRUE)

ggplot(data = tugData, aes(x = measurement, y = tug, group = subject)) +
  geom_line() +
  labs(title = "Two measurements of Timed Up and Go",
       x = "Measurement", y = "Duration (s)")
ggsave("lineplot_tug1Andtug2.png", width = 8, height = 6)
```


```{r}
tugData <- melt(capacity[, c("subject", "tug1", "tug2")], #[, (diag.vars), with = FALSE]
       id.vars = c("subject"),
       measure.vars = c("tug1", "tug2"),
       value.name = "tug",
       variable.name = "measurement",
       na.rm = TRUE)

ggplot(data = tugData, aes(x = measurement, y = tug, group = subject)) +
  geom_line() +
  labs(title = "Two measurements of Timed Up and Go",
       x = "Measurement", y = "Duration (s)")
```


```{r}
summary(rowMeans(capacity[, c("tug1", "tug2")]))
```


```{r}
hist(driver$egresstime)
```


9. scatter plot of one leg balance time vs egree and ingress time
10. histogram of subjective rating
11. scatter plot of ingress time vs egress time, gradient by subjective rating


12. The amount of missingness
```{r}
missingProp <- function(x){
  n <- length(x)
  nmiss <- sum(is.na(x))
  nmiss / n
}

driverMissProp <- apply(driver, 2, missingProp)
driverMissProp[driverMissProp>0]

capacityMissProp <- apply(capacity, 2, missingProp)
capacityMissProp[capacityMissProp>0]
```


13. BMI
```{r}
driver$BMI <- with(driver, weight.kg./(stature..without.shoes.mm./1000)^2)
summary(driver$BMI)
```

